---
title: "Brain Expression Signatures in Alzheimer's Disease"
author: "Dr Stephen Newhouse"
date: "03/02/2015"
output: html_document
---

***********************************

## getGEO

```{r load_libs, eval=FALSE}
################################################################
#  Get Data
source("http://bioconductor.org/biocLite.R")
biocLite()
biocLite(c("Biobase","limma","GEOquery"), dependencies=TRUE)
library(Biobase)
library(GEOquery)
library(limma)

################
# set working dir
setwd("/media/sjnewhouse/D/ad_brain_expression/")
```


## Array Express Data 

E-GEOD-48350  
E-GEOD-53890  
E-GEOD-29378  
E-GEOD-44772  
E-GEOD-45596  
E-GEOD-13214  
E-GEOD-36980  
E-GEOD-28146  
E-GEOD-26927  
E-GEOD-1297  
E-GEOD-6834  
E-GEOD-5281  
 
*Getting the data in one place...*

**********

#### GSE15222
#### Genetic Control of Human Brain Transcript Expression in Alzheimer Disease  
- http://www.sciencedirect.com/science/article/pii/S0002929709001086  
- Webster JA, Gibbs JR, Clarke J, Ray M et al. Genetic control of human brain transcript expression in Alzheimer disease. Am J Hum Genet 2009 Apr;84(4):445-58. PMID: 19361613  

**ABSTRACT**  
> We recently surveyed the relationship between the human brain transcriptome and genome in a series of neuropathologically normal postmortem samples. We have now analyzed additional samples with a confirmed pathologic diagnosis of late-onset Alzheimer disease (LOAD; final n = 188 controls, 176 cases). Nine percent of the cortical transcripts that we analyzed had expression profiles correlated with their genotypes in the combined cohort, and approximately 5% of transcripts had SNP-transcript relationships that could distinguish LOAD samples. Two of these transcripts have been previously implicated in LOAD candidate-gene SNP-expression screens. This study shows how the relationship between common inherited genetic variants and brain transcript expression can be used in the study of human brain disorders. We suggest that studying the transcriptome as a quantitative endo-phenotype has greater power for discovering risk SNPs influencing expression than the use of discrete diagnostic categories such as presence or absence of disease.


```{r getGEO_GSE15222, eval=FALSE}
###############################################
# load series and platform data from GEO
#
gset_GSE15222 <- getGEO("GSE15222", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset_GSE15222, file="./data/gset_GSE15222.RData",compress=TRUE)
system( paste( "echo \"./data/gset_GSE15222.RData\" >> .gitignore",sep="" ) )
rm("gset_GSE15222");gc()
# GPL2700
```

***********************************

#### GSE44772
#### Integrated Systems Approach Identifies Genetic Nodes and Networks in Late-Onset Alzheimer’s Disease

- Bin Zhang  
- http://www.sciencedirect.com/science/article/pii/S0092867413003875#  
- Zhang B, Gaiteri C, Bodea LG, Wang Z et al. Integrated systems approach identifies genetic nodes and networks in late-onset Alzheimer's disease. Cell 2013 Apr 25;153(3):707-20. PMID: 23622250  

**ABSTRACT** 
> The genetics of complex disease produce alterations in the molecular interactions of cellular pathways whose collective effect may become clear through the organized structure of molecular networks. To characterize molecular systems associated with late-onset Alzheimer’s disease (LOAD), we constructed gene-regulatory networks in 1,647 postmortem brain tissues from LOAD patients and nondemented subjects, and we demonstrate that LOAD reconfigures specific portions of the molecular interaction structure. Through an integrative network-based approach, we rank-ordered these network structures for relevance to LOAD pathology, highlighting an immune- and microglia-specific module that is dominated by genes involved in pathogen phagocytosis, contains TYROBP as a key regulator, and is upregulated in LOAD. Mouse microglia cells overexpressing intact or truncated TYROBP revealed expression changes that significantly overlapped the human brain TYROBP network. Thus the causal network structure is a useful predictor of response to gene perturbations and presents a framework to test models of disease mechanisms underlying LOAD.

Raw gene-expression data together with information related to demographics,
disease state, and technical covariates are available via the GEO
database. GEO accession numbers:-  

GSE44772  :- this is the bloody *Multi-tissue gene expression profiles of human brain*
GSE44768  : These are included in *GSE44772*  
GSE44770  : These are included in *GSE44772*    
GSE44771  : These are included in *GSE44772*  

E-GEOD-44772  
Multi-tissue gene expression profiles of human brain  
transcription profiling by array  
Homo sapiens  
690  

```{r getGEO_GSE44772, eval=FALSE}
###############################################
# load series and platform data from GEO
#
gset_GSE44772 <- getGEO("GSE44772", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset_GSE44772, file="./data/gset_GSE44772.RData",compress=TRUE)
system( paste( "echo \"./data/gset_GSE44772.RData\" >> .gitignore",sep="" ) )
rm("gset_GSE44772");gc()
# GPL4372
```

#### DONE

***********************************  

#### Accelerated neurodegeneration through chaperone-mediated oligomerization of tau
#### GSE48350 
- Blair LJ, Nordhues BA, Hill SE, Scaglione KM et al. Accelerated neurodegeneration through chaperone-mediated oligomerization of tau. J Clin Invest 2013 Oct 1;123(10):4158-69. PMID: 23999428  

*ABSTRACT*
> FK506 binding protein 51kDa (FKBP51/FKBP5) is part of a mature heat shock protein 90kDa (Hsp90) chaperone complex that preserves tau. Microarray analysis of human brains reveal that FKBP51 gene expression selectively increased with age and Alzheimer's disease, which correlated with demethylation of the regulatory regions in the FKBP5 gene. Moreover, FKBP51 levels significantly correlated with Braak pathological staging. In addition, we show that in brains devoid of FKBP51, tau levels are reduced. Recombinant FKBP51 and Hsp90 synergize to block tau clearance through the proteasome and produce T22-positive tau oligomers. Overexpression of FKBP51 in a tau transgenic mouse model revealed that FKBP51 preserved tau species, including phosphorylated and oligomeric tau that have been linked to Alzheimer's disease pathogenesis. FKBP51 blocked amyloid formation and decreased tangle load in the brain. These alterations in tau turnover and aggregate structure culminated in enhanced neurotoxicity. We propose a model where age-associated increases in FKBP51 levels can out-compete the association of other pro-degradation Hsp90 co-chaperones, resulting in neurotoxic tau accumulation. Thus, strategies aimed at attenuating FKBP51 levels or its interaction with Hsp90 could be therapeutically relevant for Alzheimer's disease and other tauopathies.
These AD cases were processed simultaneously with the control cases (young and aged) included in GSE11882

E-GEOD-48350 
Alzheimer's Disease Dataset  
transcription profiling by array  
Homo sapiens  
253   

```{r getGEO_GSE48350, eval=FALSE}
gset_GSE48350 <- getGEO("GSE48350", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset_GSE48350, file="./data/gset_GSE48350.RData",compress=TRUE)
system( paste( "echo \"./data/gset_GSE48350.RData\" >> .gitignore",sep="" ) )
rm("gset_GSE48350");gc()
```

GSE11882

*********************************** 

E-GEOD-53890  
REST and Stress Resistance in Aging and Alzheimer’s Disease  
transcription profiling by array  
Homo sapiens  
41  

```{r getGEO_GSE53890, eval=FALSE}
gset53890 <- getGEO("GSE53890", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset53890, file="./data/GSE53890.RData",compress=TRUE)
system( paste( "echo \"./data/GSE53890.RData\" >> .gitignore",sep="" ) )
```

***********************************

E-GEOD-29378  
Genes and Pathways Underlying Regional and Cell Type Changes in Alzheimer's Disease  
transcription profiling by array  
Homo sapiens  
63  

```{r getGEO_GSE29378  , eval=FALSE}
gset29378 <- getGEO("GSE29378", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset29378, file="./data/GSE29378.RData",compress=TRUE)
system( paste( "echo \"./data/GSE29378.RData\" >> .gitignore",sep="" ) )
```

***********************************

E-GEOD-45596  
Brain microvessel: Alzheimer disease vs normal  
transcription profiling by array  
Homo sapiens  
20  

***********************************

E-GEOD-13214  
Evaluation of gene expression profile in postmortem brain with Alzheimer´s disease-type neuropathological changes  
transcription profiling by array  
Homo sapiens  
128  

***********************************

E-GEOD-36980  
Expression data from post mortem Alzheimer's disease brains  
transcription profiling by array  
Homo sapiens  
79  

***********************************

E-GEOD-28146  
Microarray analyses of laser-captured hippocampus reveal distinct gray and white matter signatures associated with incipient Alzheimer’s disease  
transcription profiling by array  
Homo sapiens  
30  

***********************************

E-MTAB-656  
Alternative splicing in human temporal cortex  
transcription profiling by array  
Homo sapiens  
25  

***********************************

E-GEOD-26927  
Common neuroinflammatory pathways in neurodegenerative diseases.  
transcription profiling by array  
Homo sapiens  
118  

***********************************

E-MEXP-2280  
Transcription profiling of patients with four neurodegenerative disorders distinguishes tauopathies and identifies shared molecular pathways  
transcription profiling by array  
Homo sapiens  
31  

***********************************

E-GEOD-1297  
Transcription profiling by array of hippocampus from patients with Alzheimer's disease of varying severity
transcription profiling by array  
Homo sapiens  
31  

***********************************

E-GEOD-6834  
Temporal lobe epilepsy and Alzheimer's disease  
transcription profiling by array  
Homo sapiens  
60  

This SuperSeries is composed of the following subset Series: GSE6771: Temporal Cortex Control (mesial temporal lobe epilepsy control) GSE6773: Temporal neocortex mesial temporal lobe epilepsy GSE6774: Temporal Cortex Control (Alzheimer's disease control) GSE6775: Temporal Cortex Alzheimer's Disease GSE6777: Cerebellum Alzheimer's Disease GSE6778: Cerebellum Control (Alzheimer's disease control) Keywords: SuperSeries Refer to individual Series

***********************************

E-GEOD-5281  
Alzheimer's disease and the normal aged brain (steph-affy-human-433773)  
transcription profiling by array  
Homo sapiens  
161  

***********************************
