---
title: "Brain Expression Signatures in Alzheimer's Disease"
author: "Dr Stephen Newhouse"
date: "03/02/2015"
output: html_document
---

***********************************

## getGEO

Getting the data in one place...

**********

#### GSE15222
#### Genetic Control of Human Brain Transcript Expression in Alzheimer Disease  
- http://www.sciencedirect.com/science/article/pii/S0002929709001086  

**ABSTRACT**  

> We recently surveyed the relationship between the human brain transcriptome and genome in a series of neuropathologically normal postmortem samples. We have now analyzed additional samples with a confirmed pathologic diagnosis of late-onset Alzheimer disease (LOAD; final n = 188 controls, 176 cases). Nine percent of the cortical transcripts that we analyzed had expression profiles correlated with their genotypes in the combined cohort, and approximately 5% of transcripts had SNP-transcript relationships that could distinguish LOAD samples. Two of these transcripts have been previously implicated in LOAD candidate-gene SNP-expression screens. This study shows how the relationship between common inherited genetic variants and brain transcript expression can be used in the study of human brain disorders. We suggest that studying the transcriptome as a quantitative endo-phenotype has greater power for discovering risk SNPs influencing expression than the use of discrete diagnostic categories such as presence or absence of disease.

```{r load_libs, eval=FALSE}
################################################################
#  Get Data
library(Biobase)
library(GEOquery)
library(limma)

################
# set working dir
setwd("/media/sjnewhouse/D/ad_brain_expression/")
```


```{r getGEO_GSE15222, eval=FALSE}
###############################################
# load series and platform data from GEO
#
gset <- getGEO("GSE15222", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)

if (length(gset) > 1) idx <- grep("GPL2700", attr(gset, "names")) else idx <- 1
gset <- gset[[idx]]

#################
# save
save(gset, file="./data/GSE15222_webster2009.RData",compress=TRUE)

#######################
# add to .gitignore
system( paste( "echo \"./data/GSE15222_webster2009.RData\" >> .gitignore",sep="" ) )
###system( paste( "git rm --cached ./data/GSE15222_webster2009.RData",sep="" ) )

```

***********************************

## Array Express Data 

E-GEOD-48350  
E-GEOD-53890  
E-GEOD-29378  
E-GEOD-44772  
E-GEOD-45596  
E-GEOD-13214  
E-GEOD-36980  
E-GEOD-28146  
E-GEOD-26927  
E-GEOD-1297  
E-GEOD-6834  
E-GEOD-5281  
 

***********************************  

E-GEOD-48350 
Alzheimer's Disease Dataset  
transcription profiling by array  
Homo sapiens  
253   

```{r getGEO_GSE48350, eval=FALSE}
gset48350 <- getGEO("GSE48350", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset48350, file="./data/GSE48350.RData",compress=TRUE)
system( paste( "echo \"./data/GSE48350.RData\" >> .gitignore",sep="" ) )
```

*********************************** 

E-GEOD-53890  
REST and Stress Resistance in Aging and Alzheimer’s Disease  
transcription profiling by array  
Homo sapiens  
41  

```{r getGEO_GSE53890, eval=FALSE}
gset53890 <- getGEO("GSE53890", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset53890, file="./data/GSE53890.RData",compress=TRUE)
system( paste( "echo \"./data/GSE53890.RData\" >> .gitignore",sep="" ) )
```

***********************************

E-GEOD-29378  
Genes and Pathways Underlying Regional and Cell Type Changes in Alzheimer's Disease  
transcription profiling by array  
Homo sapiens  
63  

```{r getGEO_GSE29378  , eval=FALSE}
gset29378 <- getGEO("GSE29378", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset29378, file="./data/GSE29378.RData",compress=TRUE)
system( paste( "echo \"./data/GSE29378.RData\" >> .gitignore",sep="" ) )
```

***********************************

E-GEOD-44772  
Multi-tissue gene expression profiles of human brain  
transcription profiling by array  
Homo sapiens  
690  

```{r getGEO_GSE44772  , eval=FALSE}
gset44772<- getGEO("GSE44772", GSEMatrix =TRUE,AnnotGPL=TRUE,getGPL=TRUE)
save(gset44772, file="./data/GSE44772 .RData",compress=TRUE)
system( paste( "echo \"./data/GSE44772 .RData\" >> .gitignore",sep="" ) )
```

***********************************

E-GEOD-45596  
Brain microvessel: Alzheimer disease vs normal  
transcription profiling by array  
Homo sapiens  
20  

***********************************

E-GEOD-13214  
Evaluation of gene expression profile in postmortem brain with Alzheimer´s disease-type neuropathological changes  
transcription profiling by array  
Homo sapiens  
128  

***********************************

E-GEOD-36980  
Expression data from post mortem Alzheimer's disease brains  
transcription profiling by array  
Homo sapiens  
79  

***********************************

E-GEOD-28146  
Microarray analyses of laser-captured hippocampus reveal distinct gray and white matter signatures associated with incipient Alzheimer’s disease  
transcription profiling by array  
Homo sapiens  
30  

***********************************

E-MTAB-656  
Alternative splicing in human temporal cortex  
transcription profiling by array  
Homo sapiens  
25  

***********************************

E-GEOD-26927  
Common neuroinflammatory pathways in neurodegenerative diseases.  
transcription profiling by array  
Homo sapiens  
118  

***********************************

E-MEXP-2280  
Transcription profiling of patients with four neurodegenerative disorders distinguishes tauopathies and identifies shared molecular pathways  
transcription profiling by array  
Homo sapiens  
31  

***********************************

E-GEOD-1297  
Transcription profiling by array of hippocampus from patients with Alzheimer's disease of varying severity
transcription profiling by array  
Homo sapiens  
31  

***********************************

E-GEOD-6834  
Temporal lobe epilepsy and Alzheimer's disease  
transcription profiling by array  
Homo sapiens  
60  

This SuperSeries is composed of the following subset Series: GSE6771: Temporal Cortex Control (mesial temporal lobe epilepsy control) GSE6773: Temporal neocortex mesial temporal lobe epilepsy GSE6774: Temporal Cortex Control (Alzheimer's disease control) GSE6775: Temporal Cortex Alzheimer's Disease GSE6777: Cerebellum Alzheimer's Disease GSE6778: Cerebellum Control (Alzheimer's disease control) Keywords: SuperSeries Refer to individual Series

***********************************

E-GEOD-5281  
Alzheimer's disease and the normal aged brain (steph-affy-human-433773)  
transcription profiling by array  
Homo sapiens  
161  

***********************************
